---
title: "1713314_PDA"
author: "Bartosz Napieralski"
date: "4/4/2022"
output: html_document
---

# Data
The data that our group is going to be using can be found here:
https://www.kaggle.com/datasets/andrewmvd/heart-failure-clinical-data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("ggpubr")
#install.packages("corrplot")

library(ggplot2)
library(tidyverse)
library(ggpubr)
library(corrplot)
```


## Reading and preparing the Dataset
```{r}
raw <- read.csv("heart_failure_clinical_records_dataset.csv")
#Looking for any missing data
table(is.na(raw))

#Removing any missing columns
data <- na.omit(raw)

#Quick summary of the data
summary(data)

```
## METADATA
Here is a breakdown of all of the :

- Age - Age
- Anaemia - Decrease of red blood cells or hemoglobin (0 = No, 1 = Yes)
- Creatinine_phosphokinase -  Level of the CPK enzyme in the blood
Normal Values range between 10 to 120 mcg/L
https://www.mountsinai.org/health-library/tests/creatine-phosphokinase-test

- Diabetes - If the patient has diabetes (0 = No, 1 = Yes)
- Ejection_fraction - Percentage of blood leaving the heart at each contraction (percentage)
Normal Values have a percentage of 40% EF or higher, abnormal is below 40
https://my.clevelandclinic.org/health/articles/16950-ejection-fraction

- High_blood_pressure - If the patient has hypertension (0 = No, 1 = Yes)
- Platelets - Platelets in the blood (kiloplatelets/mL)
Normal amount of platelets is between 150,000 and 450,000 kiloplatelets/mL
https://www.hopkinsmedicine.org/health/conditions-and-diseases/what-are-platelets-and-why-are-they-important#:~:text=A%20normal%20platelet%20count%20ranges,complete%20blood%20count%20(CBC).

- Serum_creatinine - Level of serum creatinine in the blood (mg/dL)
For men it is 0.74-1.35mg/dL
For wormen it is 0.59-1.04mg/dL
https://www.mayoclinic.org/tests-procedures/creatinine-test/about/pac-20384646#:~:text=Serum%20creatinine%20level&text=The%20amount%20of%20creatinine%20in,blood%20(micromoles%2FL).

- Serum_sodium - Level of serum sodium in the blood (mEq/L)
Normal levels are between 135-145 mEq/L
https://www.mayoclinic.org/diseases-conditions/hyponatremia/symptoms-causes/syc-20373711#:~:text=A%20normal%20blood%20sodium%20level,Certain%20medications.

- Sex - Woman (0) or Man (1)
- smoking - If the patient smokes or not (0 = No, 1 = Yes)
- time - Follow up period (days)
- DEATH_EVENT - If the patient deceased during the follow up period (0 = No, 1 = Yes)



```{r}
#Adding ID to keep track of data
data$id <- seq.int(nrow(data))
data <- data %>% relocate(id)


#Assigning Character Value to correct parameters
data$sexCH <- ifelse(data$sex == 1, "Male", "Female")
data$sex <- as.logical(data$sex)
data$anaemiaCH <- ifelse(data$anaemia == 1, "Yes", "No")
data$anaemia <- as.logical(data$anaemia)
data$diabetesCH <- ifelse(data$diabetes == 1, "Yes", "No")
data$diabetes <- as.logical(data$diabetes)
data$high_blood_pressureCH <- ifelse(data$high_blood_pressure == 1, "Yes", "No")
data$high_blood_pressure <- as.logical(data$high_blood_pressure)
data$smokingCH <- ifelse(data$smoking == 1, "Yes", "No")
data$smoking <- as.logical(data$smoking)
data$DEATH_EVENT <- as.logical(data$DEATH_EVENT)


summary(data)

write.csv(data, "./data.csv", row.names = FALSE)
```

# Data Exploration

```{r}
# Since I am using large amounts of data, we can look at the population of the data

# Sex
plotSex <- ggplot(data, aes(x = sexCH)) + geom_bar(fill = "turquoise") + labs(x = "Sex", title = "Population Distributions") + theme_minimal(base_size = 10)
plotSmoke <- ggplot(data, aes(x = smokingCH)) + geom_bar(fill = "orangered") + labs(x = "Smoking") +theme_minimal(base_size = 10)
plotDiab <- ggplot(data, aes(x = diabetesCH)) + geom_bar(fill = "red") + labs(x = "Diabetic") +theme_minimal(base_size = 10)
plotDeath <- ggplot(data, aes(x = DEATH_EVENT)) + geom_bar(fill = "slateblue") + labs(x = "Died during follow up period") +theme_minimal(base_size = 10)

ggarrange(plotSex, plotSmoke, plotDiab, plotDeath)

ggplot(data, aes(x = age, y = smokingCH, color = sexCH)) + geom_boxplot()

```
```{r}
#AGE
ggplot(data, aes(x=age)) + geom_histogram(binwidth = 5, colour = "white", fill="seagreen", alpha = 0.7)+ theme_minimal(base_size = 10) + labs(title = "Age Distribution") + geom_density(eval(bquote(aes(y=..count..*5))),colour="black")
#CPK
ggplot(data, aes(x = creatinine_phosphokinase)) + geom_histogram(binwidth = 100, colour = "white", fill = "orangered", alpha = 0.7) + labs(title = "Creatinine Phosphokinase Distribution") + geom_density(eval(bquote(aes(y=..count.. * 100))),colour="black") + geom_vline(xintercept = 10, linetype = "dashed") + geom_vline(xintercept = 120, linetype = "dashed")

#Ejection Fraction
ggplot(data, aes(x = ejection_fraction )) + geom_histogram(binwidth = 5, colour =  "white", fill = "blue", alpha = 0.7) + labs(title = "Histogram of Ejection Fraction") + geom_density(eval(bquote(aes(y=..count..*5))),colour="black") + geom_vline(xintercept = 40, linetype = "dashed") + annotate("text", x = 20, y = 60, label = "Abnormal") + annotate("text", x = 55, y = 60, label = "Normal")

#Platelets
ggplot(data, aes(x = platelets)) + geom_histogram(binwidth = 25000, colour =  "white", fill = "red", alpha = 0.7) + labs(title = "Histogram of Platelet Count") + geom_density(eval(bquote(aes(y=..count.. *25000))),colour="black") + geom_vline(xintercept = 150000, linetype = "dashed") + geom_vline(xintercept = 450000, linetype = "dashed")

#Serum Creatinine
ggplot(data, aes(x = serum_creatinine)) + geom_histogram(binwidth = 0.2, colour =  "white", fill = "plum", alpha = 0.7) + labs(title = "Histogram of Creatinine in Blood") + geom_density(eval(bquote(aes(y=..count..*0.2))),colour="black") + geom_vline(xintercept = 0.59, linetype = "dashed") + geom_vline(xintercept = 1.35, linetype = "dashed")

#Serum Sodium
ggplot(data, aes(x = serum_sodium)) + geom_histogram(binwidth = 1, colour =  "white", fill = "slateblue", alpha = 0.7) + labs(title = "Histogram of Sodium in Blood") + geom_density(eval(bquote(aes(y=..count..))),colour="black") + geom_vline(xintercept = 135, linetype = "dashed") + geom_vline(xintercept = 145, linetype = "dashed")
```
 
```{r}
ggplot(data, aes(x = ejection_fraction, y = age, color = DEATH_EVENT)) + geom_point() + labs(title = "Age vs Ejection Fraction by Death Event")
```
 
 
 
 ## Correlations
 We can look for different correlation between the different ariabels to check for any different 
 
```{r}
corrplot(cor(data[2:14]), type = "lower", order = "FPC", 
         tl.col = "black", tl.srt = 30)
```
When looking at the current data we can see that there are a few different correlation are seem to arise, both positive and negative. I will investigate them further.

```{r}


cor.test(data$serum_creatinine, data$age)
#Being older does not mean that the number of creatinine in your blood specifically increases, however there is a weak correlation of 0.159
#Due to p-value being higher than 0.005 it is significant enough.
```
## Splitting the data for training/testing

```{r}
set.seed(64)
split <- sample(c(rep(0, 0.7 * nrow(data)),
                  rep(1, 0.3 * nrow(data))))

data_train <- data[split == 0,]
data_test <- data[split == 1,]

```
# Regression Model
## Creating a General Linear Model

In order for me to create a model that is able to predict possible death during the follow up period, I am going to have to correctly train a model that is able to predict the event. As such I will use a regression model to be able to predict possible outcomes.


```{r}
model <- glm(data = data_train, family = "binomial", formula = DEATH_EVENT ~ age + anaemia + creatinine_phosphokinase + diabetes + ejection_fraction + high_blood_pressure + platelets + serum_creatinine + serum_sodium + sex + smoking + time)

summary.glm(model)

simple <- step(model)

summary.glm(simple)

exp(coef(simple))

plot(simple)
```

## Checking the performance of the model:
```{r}

predictions <- predict(simple, data_test, type = "response")
output <- ifelse(predictions>0.5, 1,0)
print("Output of prediction based on test data")
test_output <- table(Actuals = data_test$DEATH_EVENT, Predictions = predictions>0.5)
print(test_output)
test_acc <- (test_output[1:1] + test_output[2:2])/nrow(data_test)

print(paste("Accuracy of the model on the test set:", test_acc))

```